---
title: "CIROQUO"
subtitle: "libKriging - Install & basics"
author: "Y. Richet, P. Havé, Y. Deville"
date: "04/2022"
output:
  ioslides_presentation:
    smaller: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## Reference

Main **entry point**:

https://github.com/libKriging/libKriging

<hr/>

For **OS** specific issues:

https://github.com/libKriging/libKriging/tree/master/docs/dev/envs


## Comaptibility matrix

<!-- ✔ ⌛️ ✘ -->
|        | Linux Ubuntu:20                             | macOS 10 & 11 (x86-64)                      | macOS 12 (ARM)**                       | Windows 10                                                                  |
|:-------|:--------------------------------------------|:--------------------------------------------|:---------------------------------------|:----------------------------------------------------------------------------|
| Python | <span style="color:green">✔</span> 3.6-3.10 | <span style="color:green">✔</span> 3.6-3.10 | <span style="color:green">✔</span> 3.9 | <span style="color:green">✔</span> 3.6-3.9                                  |
| R      | <span style="color:green">✔</span> 3.6-4.1  | <span style="color:green">✔</span> 3.6-4.1  |                                        | <span style="color:green">✔</span> 3.6-4.1                                  |
| Octave | <span style="color:green">✔</span> 5.2.0    | <span style="color:green">✔</span> 6.2      | <span style="color:green">✔</span> 6.4 | <span style="color:green">✔</span> 5.2, <span style="color:red">✘</span>6.2 |
| Matlab | <span style="color:green">️✔</span> R2022   | <span style="color:green">✔</span> R2022**  | <span style="color:red">✘</span> R2022 | <span style="color:green">✔</span> R2022**                                  |

'*': requires extra DLL
'**': no pre-built packages


## Get it

1. Go to: https://github.com/libKriging/libKriging/releases
2. Get latest package:
  * R:
    * Windows: **r**libkriging_0.5.0.**zip**
    * Linux: **r**libkriging_0.5.0_Linux-x86_64.**tgz**
    * macOS: **r**libkriging_0.5.0_macOS10.15.7-x86_64.**tgz**
  * Matlab/Octave:
    * Windows: **m**libKriging-for-**matlab**_0.5.0_**Windows10**.zip / **m**LibKriging_0.5.0_**MINGW64**_NT10.0-x86_64.tgz
    * Linux: **m**LibKriging_0.5.0_**Linux**-x86_64.tgz
    * macOS: **m**LibKriging_0.5.0_**macOS**10.15.7-x86_64.tgz 
  * Python: 
    * Windows: **py**libkriging-0.5.0-cp**XX**-**win**_amd64.whl
    * Linux: **py**libkriging-0.5.0-cp**XX**-**manylinux**_*.whl
    * macOS: **py**libkriging-0.5.0-cp**XX**-**macosx**_10_15_x86_64.whl
  * Otherwise: http://hub.irsn.cloud/ciroquo (Linux/R, Python, Matlab)
  

## Install

```{r}
install.packages("Rcpp")
install.packages(list.files(pattern = "rlibkriging_0.5.0(.*)"), repos=NULL)
```

```{octave}
# decompress dowloaded file, then
addpath("path/to/mlibkriging")
```


```{bash}
#pip3 install pylibkriging-0.5.0*.whl
immport subprocess
subprocess.check_call([sys.executable, "-m", "pip", "install", "pylibkriging-0.5.0*.whl"])
```


Please, fill survey

## Test: build model

```{r}
X <- as.matrix(c(0.0, 0.25, 0.5, 0.8, 1.0))
f <- function(x) 1 - 1 / 2 * (sin(12 * x) / (1 + x) + 2 * cos(7 * x) * x^5 + 0.7)
y <- f(X)

library(rlibkriging)
k_R <- Kriging(y, X, "gauss")
print(k_R)
```

```{octave}
X = [0.0;0.25;0.5;0.75;1.0];
f = @(x) 1-1/2.*(sin(12*x)./(1+x)+2*cos(7.*x).*x.^5+0.7)
y = f(X);
k_m = Kriging(y, X, "gauss");
disp(k_m.summary());
```

```{python}
import numpy as np
X = [0.0, 0.25, 0.5, 0.8, 1.0]
f = lambda x: (1 - 1 / 2 * (np.sin(12 * x) / (1 + x) + 2 * np.cos(7 * x) * x ** 5 + 0.7))
y = [f(xi) for xi in X]

import pylibkriging as lk
k_py = lk.Kriging(y, X, "gauss")
print(k_py.summary())
```

## Test: model predict

```{r}
...
x <- as.matrix(seq(0, 1, , 100))
p <- predict(k_R, x, TRUE, FALSE)

plot(f); points(X, y)
lines(x, p$mean, col = 'blue')
polygon(c(x, rev(x)), c(p$mean - 2 * p$stdev, rev(p$mean + 2 * p$stdev)), 
        border = NA, col = rgb(0, 0, 1, 0.2))
```

```{octave}
...
x = reshape(0:(1/99):1,100,1);
[p_mean, p_stdev] = k_m.predict(x, true, false);

h = figure(1); hold on;
plot(x,f(x)); scatter(X,f(X));
plot(x,p_mean,'b')
poly = fill([x; flip(x)], [(p_mean-2*p_stdev); flip(p_mean+2*p_stdev)],
            'b');
set( poly, 'facealpha', 0.2); hold off;
```

```{python}
...
x = np.arange(0, 1, 1 / 99)
p = k_py.predict(x, True, False)
p = {"mean": p[0], "stdev": p[1], "cov": p[2]} 

import matplotlib.pyplot as pyplot; pyplot.figure(1)
pyplot.plot(x, [f(xi) for xi in x]); pyplot.scatter(X, [f(xi) for xi in X])
pyplot.plot(x, p['mean'], color='blue')
pyplot.fill(np.concatenate((x, np.flip(x))),
            np.concatenate((p['mean'] - 2 * p['stdev'], np.flip(p['mean'] + 2 * p['stdev']))),
            color='blue', alpha=0.2)
pyplot.show()
```

## Test: model predict

```{r eval=TRUE, echo=FALSE, warning=FALSE}
X <- as.matrix(c(0.0, 0.25, 0.5, 0.8, 1.0))
f <- function(x) 1 - 1 / 2 * (sin(12 * x) / (1 + x) + 2 * cos(7 * x) * x^5 + 0.7)
y <- f(X)

library(rlibkriging)
k_R <- Kriging(y, X, "gauss")
print(k_R)

x <- as.matrix(seq(0, 1, , 100))
p <- predict(k_R, x, TRUE, FALSE)

plot(f); points(X, y)
lines(x, p$mean, col = 'blue')
polygon(c(x, rev(x)), c(p$mean - 2 * p$stdev, rev(p$mean + 2 * p$stdev)), 
        border = NA, col = rgb(0, 0, 1, 0.2))
```

## Test: model simulate


```{r}
...
s <- simulate(k_R,nsim = 10, seed = 123, x=x)

plot(f); points(X,y)
matplot(x,s,col=rgb(0,0,1,0.2),type='l',lty=1,add=T)
```

```{octave}
...
s = k_m.simulate(int32(10),int32(123), x);

h = figure(2); hold on;
plot(x,f(x)); scatter(X,f(X));
for i=1:10
   plot(x,s(:,i),'b');
end
hold off;
```

```{python}
...
s = k_py.simulate(10, 123, x)

pyplot.figure(2)
pyplot.plot(x, [f(xi) for xi in x]); pyplot.scatter(X, [f(xi) for xi in X])
for i in range(10):
    pyplot.plot(x, s[:, i], color='blue', alpha=0.2)
pyplot.show()
```


## Test: fit objective


```{r}
...
ll = function(t) logLikelihood(k_R,t)$logLikelihood
plot( ll )
```

```{octave}
...
function llt = ll (tt) 
  global k_m; 
  llt = k_m.logLikelihood(tt)
endfunction
t=0:(1/99):1
plot(t,arrayfun(@ll,t))
```

```{python}
...
def ll(t): return k_py.logLikelihood(t,False,False)[0]
t = np.arange(0,1,1/99)

pyplot.figure(3)
pyplot.plot(t, [ ll(ti) for ti in t])
pyplot.show()
```

## Fill survey :)


